---
title: "Module 12"
output: html_document
date: "2025-04-14"
---

## Main Functions

1.dclean function
This function aims to elimate NA errors and faulty data. replacing the value to zero or an average of the data points.
Example Input: 5,7,NA, 2,NA,9
Example Output: 5,7,0,2,0,9
```{r }

dclean <- function(df, replace = NA, exchange = input){
  
   
   for(colname in names(df)) {
   column <- df[[colname]]
   
   if(is.na(replace)) {
     input <- is.na(column)
   } else{
     input <- column == replace
   }
   
   if(exchange == "AVG") {
     numeric_change <- as.numeric(column)
     output <- mean(numeric_change, na.rm = TRUE)
   }else if(exchange == "MED"){
     numeric_change <- as.numeric(column)
     output <- median(numeric_change, na.rm = TRUE)
   
   } else { 
      output <- exchange 
      }
   column[input] <- output
   df[[colname]] <- column
  
}

return(df)   
   
}

air <- airquality


air <- dclean(air, replace = NA, exchange = "AVG")

```


2. dformat function
   This allows the to change the data format to something R can read and calulate with. This includes the date format and numbers separated by commas. 
Example Input: 01/21/2000
Example Output: 2000-21-01
```{r }

dformat<- function(df, format_type = input){
format_type <- as.character(format_type)
#regex patterns
date.pattern <- "\\d{1,2}/\\d{1,2}/\\d{4}"
date.pattern2 <- "\\b\\w+\\s+\\d{1,2}\\,\\s*\\d{4}\\b"
num.pattern <- "\\d{1,3},\\d{3},\\d{3}" 

    
    
   for(colname in names(df)) {
   column <- as.character(df[[colname]])
   
  if(format_type == "date") {
    dates <- grep(date.pattern, column)
    dates_value <- as.Date(column[dates], format ="%m/%d/%y")
    column[dates] <- format(dates_value, "%Y-%d-%m")
    
    dates2 <-grep(date.pattern2, column)
    dates_value2 <- as.Date(column[dates2], format ="%B %d, %Y")
    column[dates2] <- format(dates_value2, "%Y-%d-%m")
    
    }
    
   
   
   
   
  if(format_type == "number" ) {
    numbers <- grep(num.pattern, column)
  
  column[numbers] <- gsub(",","",column[numbers])

  
  }
  

    df[[colname]] <- column
    
}
   

return(df)

}


datet2 <- c("march 10, 2019", "November 10, 2000")
datet <- c("9/10/1999", "9/09/1982")
numbert <- c("1,000,000", "2,000,000")
testing <- data.frame(datet, numbert, datet2)

dformat(testing, format_type = "date")

testing
```


3.dclean.text function
This function allows the users to be able clear up any unnecessary punctuations, special character and unnecessary space.
Example Input: cash$@number
Example Output: cash number
```{r }

dclean_text <- function(vector, input){
  
  vector <- gsub(input, "" , vector, fixed = TRUE)
  vector <- gsub("[,.?/!@#$%^&`~]","",vector)
  vector <- gsub("\\s+", "", vector)
  
  
  return(vector)
}

messy <- read.csv("C:/Users/deand/Downloads/messy.csv")

messy$Email<- dclean.text(messy$Email, "COM")
```


4. dclean.names & d.rename 
These function allows allow for more control over the names of rows and columns in a dataframe. First function clean the existing names while the second rename the rows or columns the way you want to.

Example 1 input: total.NUm
Example 1 Output: total NUm
Example 2 input: toTal Numz
Example 2 Output: Total Number
```{r, }


  dclean_names <- function(df, remove) {
  if (remove == "col") {
    colnames(df) <- NULL
  } else if (remove == "row") {
    rownames(df) <- NULL
}
  return(df)
}

dclean.names(testing, "col")


  
d_rename <- function(df, x, row = FALSE, col = TRUE){
  
  if(row){
  rownames(df) <- x
  }
  if(col){
  colnames(df) <- x
  }
  return(df)
}  
namet <-c(datet, numbert, datet2)
d_rename(testing, namet, col = TRUE)
```

